---
title: "Vignette"
author: "Grace Hartley, Emma Vejcik"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ftirsr)
library(dplyr)
library(ggplot2)
```

The `ftisr` package provides easy access and user-friendly methods for researchers analyzing Biogenic Silica (BSi). (BSi) is used as a proxy for past temperatures in High Arctic settings. Typically, greater amounts of BSi in sediment cores indicate warmer temperatures. Recently paleoclimatologists have begun to use Fourier Transform Infrared (FTIR) spectroscopy to collect information on BSi. However, the offloaded relative absorbency data makes comparison with other proxies difficult. This [project](https://github.com/sds270-s22/ftirsr) is intended to develop an R package to facilitate analysis of multivariate-length samples using a model and observations in collaboration with [the Smith College PLSR SDS Capstone group](https://github.com/sds-capstone/PLSmodel). 

## Inspiration

We draw inspiration from [a previous project](https://www.causeweb.org/usproc/eusrc/2021/virtual-posters/7). Although packages that contain methods for PLSR already exist in the R universe, none provide the context and specific methods necessary for the implementation of such methods on a calibrated model specifically based on predicting BSi percentages. 


## Installation

Get the development version from GitHub:

```{r, eval = FALSE}
# If you haven't installed the remotes package yet, do so:
# install.packages("remotes")
remotes::install_github("sds270-s22/ftirsr")
```

```{r message=FALSE, warning=FALSE}
# Load package
library(ftirsr)
```


## Datasets Included
